\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-10-22 fully-abstract-comp/popl2019/paper.tex
Proof-read entire paper, making many minor changes along the way.

\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
\renewcommand{\stdreg}[3][]{\iota^{\mathrm{std},#3,#1}_{#2}}
\renewcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
\renewcommand{\codereg}[2][\mathrm{code}]{\iota^{#1}_{#2}}
\renewcommand{\staureg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta,\lrv},#2}_{#1}}

@@

Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not have been read or modified when the invoked function returns.
At the assembly level, this property also does not hold immediately.
The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
Capabilities come in different flavours.
Memory capabilities allow reading from and writing to a block of memory.
Additionally, capability machines offer some form of {\itshape object capabilities} that represent low-level encapsulated closures, i.e. a piece of code coupled with private state that it gains access to upon invocation.
The concrete mechanics of object capabilities vary between different capability machines.
% KJAA: This 'sealed' and 'common seal' business is unclear to me.
For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
Each of the two capabilities are sealed with a common seal which make them opaque.
The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015,watson_fast_2016}.

@@

In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they have been mentioned in some technical documents.}
% KJAA: Perhaps make it clear by who and when/where?
The hardware prevents these capabilities from being duplicated.
We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.

A second contribution of this work is the way in which we formulate these two properties: although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lacked evidence of generality~\cite{skorstengaard_reasoning_2017}.
We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
To accommodate this, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
As such, these components can be sure that they will only ever interact with other well-behaved components that respect our desired properties.
To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e. that components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.

Compared to~\citet{skorstengaard_reasoning_2017} that prove LSE and WBCF for a handful of examples,
this approach expresses what it means to enforce the desirable properties in a general way and makes it clear that we can support a very general class of programs.
Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
This independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.

@@

On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.

@@

In order to seal a sealable, it is necessary to have the right to do so.
The right to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$.
A set of seal gives the right to seal sealables under seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
In spirit of the memory capabilities, a set of seals has an active seal $\sigma$ which is the seal currently used to make sealed capabilities.
On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}.
When underlying capabilities are unsealed with \texttt{xjmp}, control will be transferred to one of the underlying capability which allows the capability-sealer to decide how the sealed capabilities are used.
The \texttt{xjmp} instruction will be explained in more detail later.

@@

%%% Domains
Words on \trgcm{} are capabilities and data (represented by $\ints$).
We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
We have complete memories mapping all addresses to words and memory segments mapping some addresses to words (i.e.\ partial functions).
$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.\ an out-of-bounds memory access.
An executable configuration is a register file and memory pair.

@@

\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability-related instructions.

@@

% opsem
The operational semantics of \trgcm{} is displayed in Figure~\ref{fig:target-op-sem}.
The operational semantics is defined in terms of a step relation that executes the next instruction in an executable configuration $\Phi$ which results in a new executable configuration or one of the two terminated configurations.
The executed instruction is determined by the capability in $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
If both conditions are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
The interpretations of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
The instructions that stop execution or change the flow of execution do not use $\updPcAddr{}$.
For instance, the \texttt{halt} and \texttt{fail} instructions are simply interpreted as the $\halted$ and $\failed$ configurations, respectively, and they do not use $\updPcAddr{}$.

% move w/ lin
The \texttt{move} instruction simply moves a word from one register to another.
It is, however, complicated slightly by the presence of the non-duplicable linear capabilities.
When a linear capability is moved, the source register must be cleared, so the capability is not duplicated.
This is done uniformly in the semantics using two approaches.
In the first one, we use the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
When a word $w$ is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
In the case of \texttt{move}, the source register $\rn$ is overwritten with $\linCons{\Phi(\rn)}$.
We see the other approach later.

% load w/ lin (and store)
The \texttt{store} and \texttt{load} instructions are fairly standard.
They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority, and they clear the source of a linear capability transfer.
Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
In this case, we require that the memory capability used for loading also has write-permission.

% geta, cca
The instruction \texttt{geta} projects the current address (or seal) from a capability (or set of seals), and returns $-1$ for data and sealed capabilities.
\texttt{cca} (change current address) changes the current address or seal of a capability or set of seals, respectively, by a given offset.
It handles linearity by updating the capability in-place, i.e.\ the source register is used as the target register.
This is the second approach we use to uniformly handle linearity.
The \texttt{jmp} instruction is a simple jump that just sets the $\pcreg$ register.

@@

Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
\begin{figure}[tb]
  \centering
  \[
    \begin{array}{rl c rl c rl}
    s &\in \Symbol & \phantom{make} &
    \mathrm{import} &\mathrel{::=} a \mapsfrom s & \phantom{make} &
    \mathrm{export} &\mathrel{::=} s \mapsto w\\
    \comp_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
    \comp & \multicolumn{7}{l}{\mathrel{::=} \comp_0 \mid  (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
  \end{array}
\]
\vspace{0.5cm}
\begin{mathpar}
  \inference{
    \comp_0 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
    \comp_0' = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
    \comp_0'' = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
  } {
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }
  \and
  \inference{
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }{
    (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \comp_0 \bowtie (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
  }
\end{mathpar}
% \begin{definition}[Initial execution configuration]
%   \begin{mathpar}
%     \inference{
%       c_{\mathrm{main},c} = \sealed{\sigma_1, c_{\mathrm{main},c}'} &
%       c_{\mathrm{main},d} = \sealed{\sigma_2, c_{\mathrm{main},d}'} &comp\
%       \sigma_1 = \sigma_2\\
%       \nonExec{c_{\mathrm{main},d}'}&
%       \reg(\pcreg) = c_{\mathrm{main},c}' & 
%       \reg(\rdata) = c_{\mathrm{main},d}' \\
%       \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
%       \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
%       \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0\\
%       \range{\ms_\stk} = \{0\}&
%       \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
%       [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
%       \overline{\var{import}} = \emptyset
%     }{
%       ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
%     }
%   \end{mathpar}
% \end{definition}
% \begin{definition}[Plugging a program into a context]
%   When $\var{comp'}$ is a context for component $\comp$ and $\comp' \bowtie \comp \rightsquigarrow \Phi$, 
%   then we write $\plug{\comp'}{\comp}$ for the execution configuration $\Phi$.
% \end{definition}
  \caption{Components and linking of components.}
  \label{fig:target-component-and-linking}
\end{figure}

@@

\stktokens{} is based on a traditional single stack, shared between all components.
As we are on a capability machine, it is natural to add some extra protection to stack and return pointers.
First, we replace stack pointers with stack capabilities.
When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
% KJAA: Again with the sealed pair. Again, this idea is unclear.
They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.

%informally explain how an adversary may try to abuse stack and return caps
While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
A callee that do not respect the calling convention (hence forward referred to as an adversary\footnote{The assumption on the adversary is described in detail in Section~\ref{sec:well-form-reas}.}) receives a stack capability and a return pair that they are supposed to use for the call, but the adversary can store the provided capabilities on the heap in order to use them later.
Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
In both examples our component and an adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
However, restricting the stack capability does not help when we, in the first call, give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
Generally speaking, we have no reason to ever trust a stack capability received from an untrusted component as that stack capability may have been duplicated and stored for later use.
In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
The adversary stores the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed.
The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.

% Informally explain how we prevent this using linear capabilities
As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not seem to provide sufficient guarantees to enforce LSE and WBCF.
The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code. 
Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver no longer has it or a copy of it as it is non-duplicable.} that can be used to prevent this from happening.

% Stack capability linear
The linear capabilities are put to use by requiring the stack capability to be linear.
On call, the caller splits the stack capability in two, such that they have a capability for their local stack frame and a capability for the unused part of the stack.
The stack capability for their local stack frame is sealed and used as the data part of the sealed return pair, and the stack capability for the remainder of the stack is given to the callee.
% Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
Because the stack capability is linear, the caller knows that the capability for their local stack frame cannot have an alias.
This means that an adversary would need the stack capability produced by the caller in order to access their local data.
The caller gives this capability to the adversary only in a sealed form, rendering it opaque and unusable.
This is illustrated in Figure~\ref{fig:stack-ptr-abuse-prev} and prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.

>>>>

\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
\renewcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
\renewcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
\renewcommand{\codereg}[2][\mathrm{code}]{\iota^{#1}_{#2}}
\renewcommand{\staureg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta,\lrv},#2}_{#1}}

@@

Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not have been read or modified when the invoked function returns.
At the assembly level, this property also does not hold immediately.
The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.

To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
Capabilities come in different flavours.
Memory capabilities allow reading from and writing to a block of memory.
Additionally, capability machines offer some form of {\itshape object capabilities} that represent low-level encapsulated closures, i.e. a piece of code coupled with private state that it gains access to upon invocation.
The concrete mechanics of object capabilities vary between different capability machines.
% KJAA: This 'sealed' and 'common seal' business is unclear to me.
For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
Each of the two capabilities are sealed with a common seal which make them opaque.
The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015,watson_fast_2016}.

@@

In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997} and in technical documents.
% KJAA: Perhaps make it clear by who and when/where?
The hardware prevents these capabilities from being duplicated.
We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.

A second contribution of this work is the way in which we formulate these two properties.
Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
To accommodate this, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
As such, these components can be sure that they will only ever interact with other well-behaved components that respect our desired properties.
To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e. that components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.

Compared to~\citet{skorstengaard_reasoning_2017} that prove LSE and WBCF for a handful of examples,
this approach expresses what it means to enforce the desirable properties in a general way and makes it clear that we can support a very general class of programs.
Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
Such an independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.

@@

On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015,watson_fast_2016}.

@@

In order to seal a sealable with a seal $\sigma$, it is necessary to have the right to do so.
The right to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
A set of seals is a capability that represents the authority to seal sealables with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
In spirit of memory capabilities, a set of seals has an active seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
% On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}, an operation that operates on a pair of capabilities sealed with the same seal.
The instruction will be explained in more detail below, but essentially, it unseals the pair of capabilities, transfers control to one of them (the code part of the pair) and makes the other one (the data part of the pair) available to the invoked code.

@@

%%% Domains
Words on \trgcm{} are capabilities and data (represented by integers ($\ints$)).
We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
We have complete memories mapping all addresses to words and memory segments mapping some addresses to words (i.e.\ partial functions).
$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
An executable configuration is a register file and memory pair.

@@

\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machines as well as capability-related instructions.

@@

% opsem
The operational semantics of \trgcm{} is displayed in Figure~\ref{fig:target-op-sem}.
The operational semantics is defined in terms of a step relation that executes the next instruction in an executable configuration $\Phi$ which results in a new executable configuration or one of the two terminated configurations.
The executed instruction is determined by the capability in the $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
If both conditions are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
The interpretations of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
The instructions that stop execution or change the flow of execution do not use $\updPcAddr{}$.
For instance, the \texttt{halt} and \texttt{fail} instructions are simply interpreted as the $\halted$ and $\failed$ configurations, respectively, and they do not use $\updPcAddr{}$.

% move w/ lin
The \texttt{move} instruction simply moves a word from one register to another.
It is, however, complicated slightly by the presence of the non-duplicable linear capabilities.
When a linear capability is moved, the source register must be cleared, so the capability is not duplicated.
This is done uniformly in the semantics using the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
When a word $w$ is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
In the case of \texttt{move}, the source register $\rn$ is overwritten with $\linCons{\Phi(\rn)}$.

% load w/ lin (and store)
The \texttt{store} and \texttt{load} instructions are fairly standard.
They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority.
Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
In this case, we require that the memory capability used for loading also has write-permission.

% geta, cca
The instruction \texttt{geta} projects the current address (or seal) from a capability (or set of seals), and returns $-1$ for data and sealed capabilities.
\texttt{cca} (change current address) changes the current address or seal of a capability or set of seals, respectively, by a given offset.
Note that this instruction does not need to use $\linCons{}$ like the previous ones, because it modifies the capability in-place, i.e.\ the source register is also the target register.
The \texttt{jmp} instruction is a simple jump that just sets register $\pcreg$.

@@

Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
\begin{figure}[tb]
  \centering
  \[
    \begin{array}{rl c rl c rl}
    s &\in \Symbol & \phantom{make} &
    \mathrm{import} &\mathrel{::=} a \mapsfrom s & \phantom{make} &
    \mathrm{export} &\mathrel{::=} s \mapsto w\\
    \comp_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
    \comp & \multicolumn{7}{l}{\mathrel{::=} \comp_0 \mid  (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
  \end{array}
\]
\vspace{0.5cm}
\begin{mathpar}
  \inference{
    \comp_0 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
    \comp_0' = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
    \comp_0'' = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
  } {
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }
  \and
  \inference{
    \comp_0'' = \comp_0 \bowtie \comp_0'
  }{
    (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \comp_0 \bowtie (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
  }
  \end{mathpar}
% \begin{definition}[Plugging a program into a context]
%   When $\var{comp'}$ is a context for component $\comp$ and $\comp' \bowtie \comp \rightsquigarrow \Phi$, 
%   then we write $\plug{\comp'}{\comp}$ for the execution configuration $\Phi$.
% \end{definition}
  \caption{Components and linking of components.}
  \label{fig:target-component-and-linking}
\end{figure}

@@

\stktokens{} is based on a traditional single stack, shared between all components.
To explain the technique, let us first consider how we might already add extra protection to stack and return pointers on a capability machine.
First, we replace stack pointers with stack capabilities.
When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
% KJAA: Again with the sealed pair. Again, this idea is unclear.
They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.

%informally explain how an adversary may try to abuse stack and return caps
While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
Untrusted callees receive a stack capability and a return pair that they are supposed to use for the call.
However, a malicious callee (which we will refer to as an adversary\footnote{See Section~\ref{sec:well-form-reas} for more details on our attacker model.}) can store the provided capabilities on the heap in order to use them later.
Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
In both examples our component and an adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
However, restricting the stack capability does not help when we, in the first call, give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
Generally speaking, we have no reason to ever trust a stack capability received from an untrusted component as that stack capability may have been duplicated and stored for later use.
In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
The adversary stores the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed.
The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.

% Informally explain how we prevent this using linear capabilities
As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code. 
Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver no longer has it or a copy of it as it is non-duplicable.} that can be used to prevent this from happening.

% Stack capability linear
The linear capabilities are put to use by requiring the stack capability to be linear.
On call, the caller splits the stack capability in two: one capability for their local stack frame and another one for the unused part of the stack.
The local stack frame capability is sealed and used as the data part of the sealed return pair.
The capability for the remainder of the stack is given to the callee.
% Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
Because the stack capability is linear, the caller knows that the capability for their local stack frame cannot have an alias.
This means that an adversary would need the stack capability produced by the caller in order to access their local data.
The caller gives this capability to the adversary only in a sealed form, rendering it opaque and unusable.
This is illustrated in Figure~\ref{fig:stack-ptr-abuse-prev} and prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.

% Prevention of attack 2
In a traditional calling convention with a single stack, the stack serves as a call stack keeping track of the order calls where made in and thus in which order they should be returned to.
A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
However without any enforcement, there is nothing to prevent a callee from returning from an arbitrary call on the call stack.
This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
In the presence of adversarial code, we need some mechanism to enforce that the order of the call stack is kept.
One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
The callee would have to provide this token on return to prove that it is allowed to return to the caller, and on return the token would be taken back by the caller to prevent it from being spent multiple times.
As it turns out, the stack capability for the unused part of the stack can be used as such a token in the following way:
On return the callee has to give back the stack capability they were given on invocation.
When the caller receives a stack capability back on return, they need to check that this token is actually spendable, i.e.\ check whether their stack frame is on top of the call stack or not.
They do this by attempting to restore the stack capability from before the call by splicing the return token with the stack capability for the local stack frame which at this point has been unsealed again.
If the splice is successful, then the caller knows that the two capabilities are adjacent. On the other hand, if the splice fails, then they are alerted to the fact that their stack frame may not be the topmost.
\stktokens{} uses this approach; and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}, it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
\begin{figure}
  \centering
  \begin{subfigure}{0.4\linewidth}
    \centering
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
      % recurrent parts
      \draw (-0.5,13) node {\footnotesize 0};
      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
      \stdstackstart[13]
      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
